# Respiratory Diseases and Conditions Correlated to Rise in Wildfires

##Data Used

CDC Datasets

<https://data.cdc.gov/Environmental-Health-Toxicology/Daily-County-Level-PM2-5-Concentrations-2001-2019/dqwm-pbi7/about_data>

```{r}
library(readr)
County_PM_Levels <- read_csv("Daily_County-Level_PM2.5_Concentrations__2001-2019_20241023.csv")
```


After importing my first file, I made a tibble with a sample to test cleaning and reordering the data. First I split up the "date" column and made new columns for months and days

```{r}
set.seed(1234)
date_preview <- slice_sample(County_PM_Levels, n=30)
date_preview <- mutate(date_preview, year_month_day = dmy(date))
date_preview <- mutate(date_preview, month = month(year_month_day))
date_preview <- mutate(date_preview, day = day(year_month_day))
```

Then removed the old columns "date" and "year_month_day", and made a "test" object reordered the new columns to sit next to "year"

```{r}
date_preview <- select(date_preview, !c(date, year_month_day))
date_preview <- select(date_preview, year,month,day, everything())
```

Final changes committed to County_PM_Levels dataset from the code template

```{r}

```

Gpkg file with state and county IDs imported with sf package 

```{r}
counties <- st_read("counties.gpkg")
```

county and state ID columns changed from character to numeric, original removed 

```{r}
counties_numeric <- counties %>% mutate(statefp = as.numeric(statefp)) %>% mutate(countyfp = as.numeric(countyfp))
remove(counties)
```

